Hypertension and microbiome

### 2018-12-31 rrbiome.Rmd
ALPHA-MAP plot

```{r libraries}
library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)
```

@@

filter.phenotype.data <- function(pseq,
                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
    meta(pseq) %>%
        tibble::rownames_to_column(var = "rowname") %>%
        dplyr::select(rowname, included) %>% 
        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
                      PULSEPRESSURE = SYSTM - DIASM,
                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
 #       dplyr::filter(DNA_OK == 1) %>%
        stats::na.omit() %>%
        dplyr::select(-MEN, -DNA_OK) %>%
        tibble::remove_rownames() %>%
        tibble::column_to_rownames(var = "rowname")
```

@@

```{r updating phyloseq objects}
pseq.genus <- sample_data(dset.genus)
pseq.species <- sample_data(dset.species)
```

@@

Contigency table for RR mediaction. P-value for chi-squared test is
`r chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value %>% format.pval`.

```{r self vs registered medication}
table(dset.species$BP_TREAT, dset.species$ANYDRUG) %>% kable
```

@@

```{r alpha diversity}

alphadiversity <- microbiome::global(pseq_filtered, index="shannon")
dset.alpha <- merge(dset, alphadiversity, by=0, all=TRUE)

```{r alpha diversity scatterplot, fig.width=10, fig.height=4}

ggplot((dset.alpha %>%
        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP, SEX, diversities_shannon) %>%
        gather(Covariate, Value, -SEX, -diversities_shannon)), 
       aes(x=diversities_shannon, y=Value)) +
    facet_wrap(~Covariate, ncol=4, scales = "free") +
    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
    stat_poly_eq(formula = "y ~ x",
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                 label.y = 0,
                 parse = TRUE,
                 size = 3) +   
    ylim(0, 200) +
    ggplot2::theme_minimal()

@@

ggplot(df.merged %>%
       select(sysm.signf.features, SYSTM, SEX) %>%
       gather(Covariate, Value, sysm.signf.features) %>%
       mutate(SYSTM = as.numeric(as.character(SYSTM)),
              Value = as.numeric(as.character(Value))),
       aes(x = asin(sqrt(Value)), y = SYSTM)) +
    facet_wrap(~Covariate, ncol=2, scales = "free") +
    geom_point(aes(color = SEX), size = 0.5) +
    scale_y_continuous(breaks=seq(0,200,40)) +
    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
    stat_poly_eq(formula = "y ~ x",
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                 label.y = 0,
                 parse = TRUE,
                 size = 3) +   
    ggplot2::theme_minimal()

```{r scp html, cache = FALSE}
system('scp rrbiome.html utu:articletwo')
```

>>>>

```{r libraries}
library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(kableExtra)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)
```

@@

filter.phenotype.data <- function(pseq,
                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
    meta(pseq) %>%
        tibble::rownames_to_column(var = "rowname") %>%
        dplyr::select(rowname, included) %>% 
        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
                      PULSEPRESSURE = SYSTM - DIASM,
                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
 #       dplyr::filter(DNA_OK == 1) %>%
        stats::na.omit() %>%
        dplyr::select(-MEN, -DNA_OK) %>%
        tibble::remove_rownames() %>%
        tibble::column_to_rownames(var = "rowname")

meta.merge.alphadiversity <- function(pseq, index = "shannon") {
    alphadiversity  <- microbiome::global(pseq, index = index)
    base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
        dplyr::rename(Sample_ID = Row.names) %>%
            dplyr::mutate(BL_AGE_GROUP = group_age(dset.alpha$BL_AGE))
}
```

@@

```{r updating phyloseq objects}
sample_data(pseq.genus) <- sample_data(dset.genus)
sample_data(pseq.species) <- sample_data(dset.species)
```

@@

Contigency table for RR mediaction.

```{r self vs registered medication}
tbl <- table(dset.species$BP_TREAT, dset.species$ANYDRUG)
colnames(tbl) <- c("ANYDRUG 0", "ANYDRUG 1")
rownames(tbl) <- c("BP_TREAT 0", "BP_TREAT 1")
tbl %>%
    kable(.) %>%
    kable_styling(full_width = F) %>%
    column_spec(1, bold = T)
```

$\chi^2$-test for medication `r format.pval(chisq.test(dset.species$BP_TREAT, dset.species$ANYDRUG, correct=FALSE)$p.value)`.

# Alpha diversity

@@

```{r alpha diversity}
dset.alpha <- meta.merge.alphadiversity(pseq.species)
```

```{r alpha-MAP scatter plots, fig.width=10, fig.height=4}
ggplot((dset.alpha, aes(x=diversities_shannon, y=MAP)) +
    facet_wrap(BL_AGE_GROUP, ncol=4, scales = "free") +
    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
    stat_poly_eq(formula = "y ~ x",
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                 label.y = 0,
                 parse = TRUE,
                 size = 3) +   
    ylim(0, 200) +
    ggplot2::theme_minimal() 
```

```{r alpha diversity scatterplot, fig.width=10, fig.height=4}
ggplot((dset.alpha %>%
        select(BL_AGE, BMI, SYSTM, DIASM, PULSEPRESSURE, MAP, SEX, diversities_shannon) %>%
        gather(Covariate, Value, -SEX, -diversities_shannon)), 
       aes(x=diversities_shannon, y=Value)) +
    facet_wrap(~Covariate, ncol=4, scales = "free") +
    geom_point(size=0.1, shape=1, aes(colour = SEX)) +
    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
    stat_poly_eq(formula = "y ~ x",
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                 label.y = 0,
                 parse = TRUE,
                 size = 3) +   
    ylim(0, 200) +
    ggplot2::theme_minimal() 
```

@@

ggplot(df.merged %>%
       select(sysm.signf.features, SYSTM, SEX) %>%
       gather(Covariate, Value, sysm.signf.features) %>%
       mutate(SYSTM = as.numeric(as.character(SYSTM)),
              Value = as.numeric(as.character(Value))),
       aes(x = asin(sqrt(Value)), y = SYSTM)) +
    facet_wrap(~Covariate, ncol=2, scales = "free") +
    geom_point(aes(color = SEX), size = 0.5) +
    scale_y_continuous(breaks=seq(0,200,40)) +
    geom_smooth(method=lm, se=TRUE, size=0.5, color = 'Black') +
    stat_poly_eq(formula = "y ~ x",
                 aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                 label.y = 0,
                 parse = TRUE,
                 size = 3) +   
    ggplot2::theme_minimal() 
```
